Embedding Dimension of Polyhedral Losses

### 2020-01-30 colt-20/embed-dim.tex
Iterating on 1d section

\begin{theorem}
  Let $\ell: \R \to \nonnegreals^{\Y}$ be a discrete loss.
  Then $\ell$ is $1$-embeddable if and only if there is an ordering $\R = \{r_1,\ldots,r_k\}$ of the reports such that the following two conditions hold, where $v(i)_y := : \ell(r_i)_y - \ell(r_{i-1})_y$:
  \begin{enumerate}
  \item For all $y\in\Y$, the sequence $\sgn(v(i)_y)$ is monotone in
    $i\in\{1,\ldots,k-1\}$,
  \item For all $i\in\{2,\ldots,k-1\}$, letting
    \begin{align*}
      \lambda^-(i) &= \min \left\{\frac{v(i)_y}{v({i+1})_y} : y\in\Y, v(i)_y, v({i+1})_y < 0\right\}
      \\
      \lambda^+(i) &= \max \left\{\frac{v(i)_y}{v({i+1})_y} : y\in\Y, v(i)_y, v({i+1})_y > 0\right\}~,
    \end{align*}
    we have $\lambda^-(i) \geq \lambda^+(i)$.
  \end{enumerate}

@@

  Moreover, when these conditions hold, the loss $L:\reals\to\reals^\Y$ embeds $\ell$ with $\varphi:\R\to\reals$, where
  \begin{align*}
    \varphi(r_i) &= \begin{cases}
      0 & i=1 \\
      \sum_{j=2}^i 1/\Lambda_j & i\in\{2,\ldots,k\}
    \end{cases}~,
    \\
    L(u) &= \begin{cases}
      \ell(r_1) - u K \ones & u \leq 0 \\
      \ell(r_i) + \Lambda_i \cdot (u - \varphi(r_i)) \cdot  (\ell(r_{i+1})-\ell(r_i)) & u \in [\varphi(r_i),\varphi(r_{i+1})] \\
      \ell(r_k) + u K \ones & u \geq \varphi(r_k)
    \end{cases}~,
  \end{align*}
  where $\Lambda_i := \prod_{j=2}^i \lambda(j)$ and $K=1+\max_{r\in\R,y\in\Y} |\ell(r)_y|$.
\end{theorem}

The main result of this section is Theorem \ref{thm:orderable-1-embed}.
\begin{theorem} \label{thm:orderable-1-embed}
  Given a discrete loss $\ell: \R \to \nonnegreals^{\Y}$, the following are equivalent:
  \begin{enumerate}
    \item The property elicited by $\ell$ is orderable.
    \item $\ell$ is $1$-embeddable.
    \item $\ell$ has some polyhedral calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$.
    \item $\ell$ has some convex calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$.
  \end{enumerate}
\end{theorem}
This theorem implies that one can directly test if a property is $1$-embeddable via the orderable criterion.
Orderability can be tested using the Bayes risk of $\ell$, a polyhedral function defined by $p \mapsto \max_{r \in \R} \inprod{p}{\ell(r)}$.
A direct approach is to project this function down to the simplex to give the level sets $\gamma_r$, as in Figure~\ref{fig:intersection-graph-ex}.
Another approach is to note that the intersection graph is directly given by finding those pairs $r,r'$ that simultaneously maximize the Bayes risk, i.e. cases where for some $p$, we have $r,r' \in \argmax_{r'' \in \R} \inprod{p}{\ell(r'')}$.

@@

The feasibility program can be viewed as a low-rank matrix problem, namely: do there exist a set of rank-$d$ matrices that are pointwise dominated by $B$, sharing the left factor $V$, whose right factors $X^j$ respectively satisfy a subspace constraint?
We will see in Section~\ref{subsec:example-abstain}\bo{in which section?} that for the important example of abstain loss, the constraints simplify into a more pure low-rank matrix problem.
In particular, for $d=n-1$ a solution always exists (e.g. \bo{cite} \cite[Theorem 2]{finocchiaro2019embedding}), found by taking the convex conjugate of the negative Bayes risk of $\ell$ for each outcome and subtracting the report $u$ \bo{and then what?} \jessie{Everything I just added is for $n$ dimensions, but I don't know where the $n-1$ dim is}.

@@

To gain a better understanding about the QFP in Definition~\ref{def:qfp}, we prove the following:
\begin{lemma} \label{lemma:D-polytope}
  For any $\T^r$, $D(\T^r)$ is a polytope (in particular, is convex).
\end{lemma}
\begin{proof}
   Recall by definition, the notation $\oplus_p \T^r = \{\sum_y p(y) x^r_y : x^r_y \in T^r_y (\forall y)\}$.
   Each $T^r_y$ is a polytope, so $p(y) T^r_y$ is a polytope.
   The Minkowski sum of polytopes is a polytope, so $\oplus_p \T^r$ is a polytope by~\cite[Section 1.2]{weibel2007minkowski}.
  \bo{TODO}\jessie{Done?}
\end{proof}
\begin{lemma}  \label{lemma:minkowski-support}
  Given polytopes $\T^r$, there exists a finite set of normal vectors $w_1,\ldots,w_K \in \reals^d$ such that, for all $p \in \simplex$, $\oplus_p \T^r = \{x : \inprod{w_i}{x} \leq \sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{w_i}{x} \}$.
\end{lemma}
\begin{proof}
   For each $p$, $\oplus_p \T^r$ is a polytope.

  \bo{TODO: it turns out that $p$ doesn't matter, only its support; and there are only finitely many possible supports (or even better, the normals that are complete for the uniform minkowski sum turn out to be complete for all other $p$-combinations).}
\end{proof}
\begin{lemma} \label{lemma:E-to-B}
  Let $C = \{p : Bp \geq \vec 0 \}$ where $B$ has the minimum possible number of rows to capture $C$, and suppose $C = \{p : Ep \geq \vec 0 \}$.
  Then for each row in $B$ there is some (unique) row in $E$ that is equal to $\alpha B$ for some positive $\alpha$.
\end{lemma}
\begin{proof}
  \bo{? Cite} \jessie{Is left as an exercise in the following text:} \cite[Exercise 2.15]{ziegler2012lectures}
\end{proof}

>>>>

\begin{theorem}
  Let $\ell: \R \to \nonnegreals^{\Y}$ be a discrete loss.
  Then $\ell$ is $1$-embeddable if and only if there is an ordering $\R = \{r_1,\ldots,r_k\}$ of the reports such that the following two conditions hold, where $v(i)_y := \ell(r_i)_y - \ell(r_{i-1})_y$:
  \begin{enumerate}
  \item For all $y\in\Y$, the sequence $\sgn(v(i)_y)$ is monotone in
    $i\in\{1,\ldots,k-1\}$,
  \item For all $i\in\{2,\ldots,k-1\}$, letting
    \begin{align*}
      \lambda^-(i) &= \min \left\{\frac{v(i)_y}{v({i+1})_y} : y\in\Y, v(i)_y, v({i+1})_y < 0\right\}
      \\
      \lambda^+(i) &= \max \left\{\frac{v(i)_y}{v({i+1})_y} : y\in\Y, v(i)_y, v({i+1})_y > 0\right\}~,
    \end{align*}
    we have $\lambda^-(i) \geq \lambda^+(i)$. \jessie{Do we need to be more careful defining these when the sets are empty?}
  \end{enumerate}

@@

  Moreover, when these conditions hold, the loss $L:\reals\to\reals^\Y$ embeds $\ell$ with $\varphi:\R\to\reals$, \jessie{Have been using $\phi$ instead of $\varphi$ for embeddings in this paper so far. I have a slight preference for $\varphi$ since it matches the last paper, but I don't care too much... just need to pick one.}where
  \begin{align*}
    \varphi(r_i) &= \begin{cases}
      0 & i=1 \\
      \sum_{j=2}^i 1/\Lambda_j & i\in\{2,\ldots,k\}
    \end{cases}~,
    \\
    L(u) &= \begin{cases}
      \ell(r_1) - u K \ones & u \leq 0 \\
      \ell(r_i) + \Lambda_i \cdot (u - \varphi(r_i)) \cdot  (\ell(r_{i+1})-\ell(r_i)) & u \in [\varphi(r_i),\varphi(r_{i+1})] \\
      \ell(r_k) + u K \ones & u \geq \varphi(r_k)
    \end{cases}~,
  \end{align*}
  where $\Lambda_i := \prod_{j=2}^i \lambda(j)$ and $K=1+\max_{r\in\R,y\in\Y} |\ell(r)_y|$.
  \jessie{In the cases: what if $\varphi(r_1) < 0$?}
\end{theorem}

The main result of this section is Theorem \ref{thm:orderable-1-embed}.
\begin{theorem} \label{thm:orderable-1-embed}
  Given a discrete loss $\ell: \R \to \nonnegreals^{\Y}$, the following are equivalent:
  \begin{enumerate}
    \item The property elicited by $\ell$ is orderable.
    \item $\ell$ is $1$-embeddable.
    \item $\ell$ has some polyhedral calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$.
    \item $\ell$ has some convex calibrated surrogate loss $L: \reals \to \nonnegreals^{\Y}$.
  \end{enumerate}
\end{theorem}
This theorem implies that one can directly test if a property is $1$-embeddable via the orderable criterion.
Orderability can be tested using the Bayes risk of $\ell$, a polyhedral function defined by $p \mapsto \max_{r \in \R} \inprod{p}{\ell(r)}$.
A direct approach is to project this function down to the simplex to give the level sets $\gamma_r$, as in Figures~\ref{fig:intersection-graph-ex} and~\ref{fig:abstain-alpha-half}.
Another approach is to note that the intersection graph is directly given by finding those pairs $r,r'$ that simultaneously maximize the Bayes risk, i.e. cases where for some $p$, we have $r,r' \in \argmax_{r'' \in \R} \inprod{p}{\ell(r'')}$.

@@

The feasibility program can be viewed as a low-rank matrix problem, namely: do there exist a set of rank-$d$ matrices that are pointwise dominated by $B$, sharing the left factor $V$, whose right factors $X^j$ respectively satisfy a subspace constraint?
We will see in Section~\ref{subsec:example-abstain}\bo{in which section?} that for the important example of abstain loss, the constraints simplify into a more pure low-rank matrix problem.
In particular, for $d=n-1$ a solution always exists (e.g. \bo{cite} \cite[Theorem 2]{finocchiaro2019embedding}), found by taking the convex conjugate of the negative Bayes risk of $\ell$ for each outcome and subtracting the report $u$ \bo{and then what?} \jessie{Everything I just added is for $n$ dimensions, but I don't know where the $n-1$ dim is... possibly~\cite{abernethy2013efficient}?}.

@@

To gain a better understanding about the QFP in Definition~\ref{def:qfp}, we prove the following:
\begin{lemma} \label{lemma:D-polytope}
  For any $\T^r$, $D(\T^r)$ is a polytope (in particular, is convex).
\end{lemma}
\begin{proof}
   Recall by definition, the notation $\oplus_p \T^r = \{\sum_y p(y) x^r_y : x^r_y \in T^r_y (\forall y)\}$.
   Each $T^r_y$ is a polytope, so $p(y) T^r_y$ is a polytope.
   The Minkowski sum of polytopes is a polytope, so $\oplus_p \T^r$ is a polytope by~\cite[Section 1.2]{weibel2007minkowski}.
  \bo{TODO}\jessie{Done?}
\end{proof}
\begin{lemma}  \label{lemma:minkowski-support}
  Given polytopes $\T^r$, there exists a finite set of normal vectors $w_1,\ldots,w_K \in \reals^d$ such that, for all $p \in \simplex$, $\oplus_p \T^r = \{x : \inprod{w_i}{x} \leq \sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{w_i}{x} \}$.
\end{lemma}
\begin{proof}
   For each $p$, $\oplus_p \T^r$ is a polytope.
  \bo{TODO: it turns out that $p$ doesn't matter, only its support; and there are only finitely many possible supports (or even better, the normals that are complete for the uniform minkowski sum turn out to be complete for all other $p$-combinations).}
\end{proof}
\begin{lemma} \label{lemma:E-to-B}
  Let $C = \{p : Bp \geq \vec 0 \}$ where $B$ has the minimum possible number of rows to capture $C$, and suppose $C = \{p : Ep \geq \vec 0 \}$.
  Then for each row in $B$ there is some (unique) row in $E$ that is equal to $\alpha B$ for some positive $\alpha$.
\end{lemma}
\begin{proof}
  \bo{? Cite} \jessie{Is left as an exercise in the following text:} \cite[Exercise 2.15]{ziegler2012lectures}
\end{proof}
